Non-word repetition in Yélî Dnye

### 2020-09-16 manuscript.tex
replied to mc comments

\begin{itemize}
\tightlist
\item
  add effect sizes to all analyses
\item
  reply to questions
\end{itemize}

@@

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether.

@@

Taking into account only the first attempts, we averaged attempts by each of the 24 children who had data for first attempts; their ages ranged from 3.90 to 10.20 (M = 6.50, SD = 1.50).

The overall NWR score was M = 60\% (SD = 15\%). Scores based on phonemes are even higher M = 77\% (SD = 10\%). The phoneme-based normalized Levenshtein distance was M = 23\% (SD = 10\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

Finally, we looked into the frequency with which mispronunciations resulted in real words. Nearly all of them were: 97\%.

\textbf{TO DISCUSS TOGETHER this is higher than any previous report! can we be sure of it?}

@@

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 59 (17) & 41 (17)\\
79 (22) & 92 (9) & 8 (9)\\
78 (19) & 93 (7) & 7 (7)\\
74 (32) & 91 (12) & 9 (12)\\
\bottomrule
\end{tabular}

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (1,421.71) = 0.38, p = 0.07). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (0.03) = 21.40, p = 0.98), birth order (data missing for 17 children, 2,652.32) = -0.15, p = 0.47) or maternal education (data missing for 2 children, 2,101.60) = 0.09, p = 0.69)

>>>>

\begin{itemize}
\tightlist
\item
  add information about proportion of mispronunciations that are real based on other papers
\item
  add effect sizes to all analyses
\item
  ? correct treatment of words whose mispronunciation is an English word:
  mispronunciation meaning
  361 two two
  362 two two
  879 no-pinêt English: no peanut
  908 nodedi English: no daddy
  933 no--dadi English: no daddy
  949 petrumi English: ``true''
  950 petrumi English: ``true''
  951 petrumi English: ``true''
  1373 wee English: where
  1374 wee English: where
  1375 wee English: where
  1379 wee English: where
  1380 wee English: where
  1388 wee English: where
  1394 wee English: where
  1396 wee English: where
  1398 wee English: where
  1399 wee English: where
  1402 wat English: what
  1405 wat English: what
  1411 wee English: where
\end{itemize}

@@

The first phase was making sure the child understood the task. This was explained orally and the first training item was presented. Often, children froze and did not say anything. If this happened, then we followed this procedure. First, the informant insisted. If the child still did not say anything, the informant asked the child to repeat a real word, and another, and another. If the child could repeat these correctly, then we provided the recorded training item over headphones again. Most children successfully started repeating the items presented over headphones at this point; a few further needed the local informant to model the behavior (i.e., they would hear the item again, and she would say it; then we would play it again, and ask the child to say it). A small minority still failed to repeat the item after hearing it over headphones. If that occurred, we tried with the second training item, at which point some children got it and could continue. A small minority, however, failed to repeat this one, as well as the third training item, in which case we stopped the test altogether (see Participants section for exclusions).

@@

Taking into account only the first attempts, we averaged attempts by each of the 42 children who had data for first attempts; their ages ranged from NA to NA (M = NA, SD = NA).

The overall NWR score was M = 65\% (SD = 15\%). Scores based on phonemes are even higher M = 78\% (SD = 9\%). The phoneme-based normalized Levenshtein distance was M = 22\% (SD = 9\%), meaning that about a fifth of phonemes were were substituted, inserted, or deleted. Notice that the normalized Levenshtein distance is the complement of phoneme-based scores.

@@

Finally, we looked into the frequency with which mispronunciations resulted in real words. Two thirds of them were: 62\%.

@@

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 59 (17) & 41 (17)\\
80 (22) & 93 (9) & 7 (9)\\
79 (19) & 93 (7) & 7 (7)\\
74 (32) & 91 (12) & 9 (12)\\
\bottomrule
\end{tabular}

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (6,496.39) = 0.43, p = 0.00). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (-0.61) = 27.83, p = 0.55), birth order (data missing for 17 children, rho = (2,652.32) = -0.15, p = 0.47) or maternal education (data missing for 2 children, rho = ( 9,022.76) = 0.09, p = 0.60).
